---
title: "Project Proposal"
author: "Team 23"
output:
  pdf_document:
    keep_tex: true
    latex_engine: pdflatex
    

    
---

```{r load-packages, message = FALSE,include=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(countrycode)
```

```{r setup, include = FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
```

## 1. Introduction

\color{blue} Kiva is a non-profit organization that provides microfinance loans to underserved communities around the world. However, not all loans are equally successful in achieving their intended impact. Our research aims to contribute to the larger problem of improving the effectiveness of microfinance by exploring the characteristics of loans that are more likely to be funded, repaid, or defaulted.


Specifically, our research question is whether we can predict the likelihood of loan default using loan characteristics available in the Kiva dataset. Additionally, we are investigating whether loans with more lenders have a higher likelihood of repayment, potentially leveraging the "wisdom of the crowd" to improve loan outcomes.


This research is important because microfinance loans can have a significant impact on the lives of borrowers and their communities, but defaults can harm not only the borrower but also the lender and the broader microfinance ecosystem. By identifying the loan characteristics that are most predictive of default and repayment, we can help lenders make more informed decisions about loan approval and investment, as well as help borrowers and organizations design more effective loan programs. Our approach involves analyzing the Kiva dataset using machine learning techniques to build models that predict loan default and assess the impact of the number of lenders on loan outcomes.

In addition to our study , numerous studies have been conducted on the Kiva platform, exploring various aspects such as borrower characteristics, loan repayment rates, and lender behavior. These studies provide valuable insights into the factors that impact loan performance and can guide our analysis. For this research, we will be basing our analysis on a couple of these studies, including "I Loan Because...:"  (available at http://www-personal.umich.edu/~qmei/pub/wsdm2012-liu.pdf). This study provides a deeper understanding of why lenders choose to invest in Kiva loans and how they prioritize their lending decisions. By incorporating these findings into our analysis, we hope to gain a more comprehensive understanding of loan funding and repayment on the Kiva platform.
\color{black}

## 2. Data

```{r,warning=FALSE,include=FALSE}
kiva_data <- read.csv("/Users/eyal.benbarouch/Documents/kiva_data/loans_splitted_0.csv")

kiva_data%>%
  select_if(is.numeric)%>%
  summary()

kiva_data <- kiva_data %>%
  mutate(funded = ifelse(STATUS=="funded",1,0))
kiva_data$funded <- as.factor(kiva_data$funded)

```



```{r}
kiva_data <- read.csv("/Users/eyal.benbarouch/Documents/שטיות שלי/Studies/סמסטר ד/תכנות וניתוח נתונים/פריוקט סיום/kiva/loans.csv")

kiva_ml_data <- read.csv("/Users/eyal.benbarouch/Documents/שטיות שלי/Studies/סמסטר ד/תכנות וניתוח נתונים/פריוקט סיום/kiva/kiva_dataset_ML.csv")
inner_join_table <- merge(kiva_ml_data,kiva_data,by.y="LOAN_ID",by.x="id",all = FALSE)
kiva_joined_data <- inner_join_table
save_path <- "/Users/eyal.benbarouch/Documents/שטיות שלי/Studies/סמסטר ד/תכנות וניתוח נתונים/פריוקט סיום/kiva/kiva_joined_data.csv"
write.csv(kiva_joined_data,save_path,row.names = FALSE)

```


```{r}
status_count <-table(kiva_joined_data$Status)
status_count
prop.table(status_count)
head(kiva_joined_data)
kiva_join_sample <- kiva_joined_data %>% sample_n(100)
save_path_sample <- "/Users/eyal.benbarouch/Documents/שטיות שלי/Studies/סמסטר ד/תכנות וניתוח נתונים/פריוקט סיום/kiva/kiva_joined_sample_data.csv"
write.csv(kiva_join_sample,save_path_sample,row.names=FALSE)

```

## 3. Preliminary results

\color{blue} In this Graph we want to get a feel of our sub question, if there is any bias in loan funding on Kiva , by checking funding counts by continent 



```{r,warning=FALSE,fig.width=4,fig.height=4}


continent_lookup <- data.frame(countrycode::codelist$iso2c, countrycode::codelist$region)


colnames(continent_lookup) <- c("country_code", "continent")

kiva_data <- kiva_data %>%
  left_join(continent_lookup, by = c("COUNTRY_CODE" = "country_code"))


  ggplot(data = kiva_data, aes(x = continent, fill = funded)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "blue"), labels = c("Not Funded", "Funded")) + 
  theme(axis.text.x=element_text(angle=45,vjust = 1,hjust = 1))
  

#axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1
```

\color{blue} In this Graph we want to get a feel of our sub question, if there the "wisdom of the crowd" factor comes into play - meaning the higher the number of lenders the lower chance of default


```{r}
num_defaulted_per_numlenders <- kiva_joined_data %>%
  group_by(NUM_LENDERS_TOTAL) %>%
  filter(Status == "defaulted") %>%
  summarise(num_defualted = n())

num_per_num_lenders <-  kiva_joined_data %>%
  group_by(NUM_LENDERS_TOTAL) %>%
  summarise(total_num_loans = n())

num_total_vs_num_defaulted <- merge(num_per_num_lenders,num_defaulted_per_numlenders)

num_total_vs_num_defaulted <- num_total_vs_num_defaulted %>%
  mutate(prop_defaulted = num_defualted/total_num_loans)

num_total_vs_num_defaulted %>%
  filter(total_num_loans > 50) %>%
  ggplot(aes(x = NUM_LENDERS_TOTAL, y = prop_defaulted)) +
  geom_line(color = "steelblue", size = 0.7 ) +
  geom_smooth(method = "loess", se = FALSE, color = "darkred", linetype = "solid") +
  labs(title = "Proportion of defaulted loans by number of lenders",
       x = "Number of lenders",
       y = "Proportion of defaulted loans") +
  theme_minimal()
  






```

As seen in the graph in the beggining the lower number of lenders the higher chacne of default but once we pass a certain point around 75 lenders the chance of default rises , it would be interesting to find out why


As seen in the graph there are contitents that seem to have higher rates of funding declines , In our project we will check to see if the reason is bias.



<!-- \color{blue} In the next two graphs we want to get a feel of sector segmentation of funding -->

```{r,warning=FALSE,fig.width=7,fig.height=5,include=FALSE}
library(scales)
ggplot(kiva_data, aes(x = SECTOR_NAME, y = FUNDED_AMOUNT)) + 
  geom_bar(stat = "summary", fun = sum,fill = "steelblue") +
  #geom_boxplot(fill = "lightblue", color = "blue") +
  theme_bw() +
  labs(x = "Sector Name", y = "Funded Amount", title = "Funded Amount by Sector Name") + 
  coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.margin = unit(c(1, 10, 1, 1), "cm"),
        plot.title = element_text(size = 20)) + 
  scale_y_continuous(labels = scales::comma) +
  theme(plot.title = element_text(hjust = 0.5))

```


<!-- \color{blue} As seen in the graph the top sectors are food and agriculture. Note that this data is not normelized yet so it may be biased. -->


In This Graph we examine the proportion of defaulted declines per sector 

```{r,warning=FALSE,fig.width=4,fig.height=3}
kiva_data <- kiva_data %>%
  mutate(funded = ifelse(STATUS=="funded",1,0))


total_per_sector <-  kiva_data %>% 
  group_by(SECTOR_NAME) %>% 
  summarise(count = n())


total_not_funded_per_sector <- kiva_data %>%
  group_by(SECTOR_NAME) %>% 
  summarise(not_funded = sum(ifelse(funded == 0, 1, 0)))


proportions <- total_not_funded_per_sector$not_funded / total_per_sector$count

proportions_tbl <- tibble(Sectors = total_not_funded_per_sector$SECTOR_NAME,
                          proportions = proportions)




ggplot(proportions_tbl, aes(x = reorder(Sectors, -proportions), y = proportions)) +
  geom_col(color = "#008080", size = 1.2) +
  labs(x = "Sector", y = "Proportion of Loans Not Funded", 
       title = "Proportion of Loans Not Funded by Sector") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


```

As seen in the Graph - more "luxrish" sectors such as housing and clothing have more chance of being declined than more "moral" sectors such as Education and Health \color{black}





## 4. Data analysis plan

### Section 4 - Work plan:

\color{blue}1. Our work plan would be to create a prediction model that would try to imply whether a loan was funded or not:

 - The outcome (Y) is going to be funded (1 means YES, 0 means NO). the predictors (X) are going to be the ethnicity variables and buisness variables. the reason for this choice is based on basic logic and nothing else - obviously the predictors can be updated during the project.

2. In order to perform our analysis, we have a large dataset that has been split into four equally sized subsets, each containing 400K rows. This allows us to create comparison groups and perform cross-validation by further splitting the subsets into test and train subsets. We will use this approach to compare and evaluate different models and methods for our analysis. By using multiple comparison groups and cross-validation, we can ensure that our results are robust and not biased by a single subset or model.

3. The methods that we would use is first of all logistic regression models, in order to find the significant predictors (predictors with p-value smaller or equal to 0.05). Second of all, we would create a new model with the predictors we found and evaluate the model in order to check our results. Nevertheless, we would normalize our data in order to prevent a bias in the result, for instance the reason most of the money founded in Agriculture is because this sector is more common in the countries that are most common in Kiva.

4. To support the hypothesized question of predicting funding of a loan using Logistic Regression, we would need several statistical results, such as:

	4.1.Model performance metrics: We need to measure the accuracy of the model's predictions, such as sensitivity, specificity, and area under the curve (AUC).
	4.2.Significance of predictor variables: We need to determine the significance of each predictor variable in the model, which can be done by examining the p-values of the coefficient estimates.

5. For our subquestion of the research - Investigate if there is any bias in loan funding on Kiva we will
need to perform statistical tests to compare the funding rates between different groups of loans. This could involve comparing the funding rates between loans from different countries, different loan purposes, different loan sizes, or any other relevant categories.

6. Teamwork: Our intentions are to set up a weekly meeting, and talk about the goals on the following week and to assign tasks to each one in order to continue more efficiently in the project. this approach would let each person to continue separately and perhaps back home.






## Appendix
https://escholarship.org/uc/item/9cc4t85b#article_main - Loan Repayment Prediction Using Machine Learning Algorithms by Chang Han 2019


### Data README

```{r include_data_readme, comment=''}
cat(readLines('../data/README.md'), sep = '\n')
```

### Source code

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
